---
title: "COVID-19 Dashboard"
output: 
  flexdashboard::flex_dashboard:
    theme: flatly
    social: menu    
    logo : "favicon-32x32.png"
    source_code: https://github.com/amannj/COVID-19
runtime: shiny    
---

```{r setup, include=FALSE}
## Time stamp  --------------
date_today <- substr(Sys.time(), 1, 10)

## Create folders  --------------
dir.create(paste0("figures/", date_today))
dir.create(paste0("data/", date_today))

## Options and packages  --------------
knitr::opts_chunk$set(
  fig.path = paste0("figures/", date_today, "/"),
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
library(lubridate)
library(tidyverse)
library(kableExtra)
library(rlang)
library(DT)
library(scales)
library(ggrepel)
library(WDI)
library(plotly)
library(ggplotlyExtra)

## Colour pattern
cp_custom <- c("#21908CFF", "#FDE725FF", "#440154FF")

## Function files
source("functions.R")
```

Global {data-icon="fa-globe"}
===================================== 

```{r download_CSSE}
# Download and store in separate folder every day  --------------
# name_files <- c(
#   "time_series_covid19_confirmed_global.csv",
#   "time_series_covid19_deaths_global.csv",
#   "time_series_covid19_recovered_global.csv"
# )
# 
# bin <- lapply(name_files, store_local)
```

```{r transform_CSSE}
# Load raw data into R   --------------
df_confirmed <- read.csv(paste0("data/CSSE/time_series_covid19_confirmed_global.csv"), stringsAsFactors = FALSE)
df_dead <- read.csv(paste0("data/CSSE/time_series_covid19_deaths_global.csv"), stringsAsFactors = FALSE)
df_recovered <- read.csv(paste0("data/CSSE/time_series_covid19_recovered_global.csv"), stringsAsFactors = FALSE)


# Create final data frames   --------------

## Merge data sets
df_confirmed %>%
  tidy_data("confirmed") %>%
  full_join(df_dead %>%
    tidy_data("dead"),
  by = c("country", "date")
  ) %>%
  full_join(df_recovered %>%
    tidy_data("recovered"),
  by = c("country", "date")
  ) %>%
  mutate(remaining = confirmed - dead - recovered) %>%
  ungroup() %>%
  mutate(iso2c = countrycode::countrycode(country, origin = "country.name", destination = "iso2c")) -> df_fin

## Get most recent WDI population data
WDI(
  country = df_fin %>% distinct(iso2c) %>% pull(),
  indicator = "SP.POP.TOTL", start = 2015
) %>%
  group_by(iso2c) %>%
  na.omit() %>%
  filter(year == max(year)) %>%
  select(iso2c, pop = "SP.POP.TOTL") -> df_pop

## Merge data
df_fin %>%
  left_join(df_pop, by = "iso2c") -> df_fin
```

Inputs {.sidebar data-width=300}
-----------------------------------------------------------------------
<br>
**Instructions. ** Using the control elements provided below to

- add/remove entries from the list of 'selected countries',
- define time span, and
- trigger data transformations
  - using linear/logarithmic scale;
  - total /per thousand inhabitants;

for interactive figures in central panel. 

***

```{r}
## Country selection -------------
selectInput("select_countries", h4("Select countries"),
  choices = df_fin %>% distinct(country) %>% pull(),
  multiple = TRUE,
  selected = c("Austria", "Germany", "Italy", "Spain", "United Kingdom", "France", "China", "Japan", "Sweden", "Switzerland")
)
```

***

```{r}
## Select data range --------
dateRangeInput(
  "select_date",
  h4("Select range"),
  start = df_fin %>% ungroup() %>% summarise(date = min(date)) %>% pull(),
  end = df_fin %>% ungroup() %>% summarise(date = max(date)) %>% pull(),
  format = "dd-mm-yyyy"
)
```

***

```{r}
## Series transformation ---------
radioButtons("transform_series",
  h4("Transform series"),
  inline = FALSE,
  c(
    "Linear" = "linear",
    "Logarithm" = "log"
  )
)

## Series scaling ---------
radioButtons("scale_series", "",
  inline = FALSE,
  c(
    "Per thousand inhabitants" = "pc",
    "Total cases" = "cases"
  )
)
```
    
Column {data-width=600, .tabset} 
-------------------------------------
    
### Cumulative cases over time

```{r}
output$cumulative <- renderPlotly({
  plot_TS(df_fin,
    country_select = reactive(input$select_countries),
    start_date = reactive(input$select_date[1]),
    end_date = reactive(input$select_date[2]),
    transform_series = reactive(input$transform_series),
    scale_series = reactive(input$scale_series),
    date_today = date_today,
    is.ts = FALSE
  ) -> p3


  # Manually set tooltip info
  plotly_build(p3)
})

plotlyOutput("cumulative", width = "100%")
```

### Progression by country

```{r}
output$ts <- renderPlotly({
  plot_TS(df_fin,
    country_select = reactive(input$select_countries),
    start_date = reactive(input$select_date[1]),
    end_date = reactive(input$select_date[2]),
    transform_series = reactive(input$transform_series),
    scale_series = reactive(input$scale_series),
    date_today = date_today,
    is.ts = TRUE
  ) -> p4


  # Manually set tooltip info
  plotly_build(p4)
})

plotlyOutput("ts", width = "100%")
```
   
Column {data-width=400, .tabset}
-------------------------------------

### Ten Most Affected (full sample)
    
```{r most.affected.all}
# Set topX to top-10  --------------
x <- 10

# Aggregate data full sample --------------

## Extract full sample topX countries
df_fin %>%
  group_by(country) %>%
  summarize_at(c("confirmed", "dead", "recovered", "remaining"), max) -> df_topX_all

# Extract list of topX countries for each category for full sample
rbind(
  extract_topX(df_topX_all, X = x, "confirmed") %>% select(country, series),
  extract_topX(df_topX_all, X = x, "dead") %>% select(country, series),
  extract_topX(df_topX_all, X = x, "recovered") %>% select(country, series),
  extract_topX(df_topX_all %>% filter(country != "US"), X = x, "remaining") %>% select(country, series)
) %>% as_tibble() -> topX_countries_all

## Plot full sample top countries
rbind(
  extract_topX(df_topX_all, X = x, "confirmed"),
  extract_topX(df_topX_all, X = x, "dead"),
  extract_topX(df_topX_all, X = x, "recovered"),
  extract_topX(df_topX_all %>% filter(country != "US"), X = x, "remaining")
) %>%
  ggplot(aes(x = country, y = `share (%)`, fill = country, cases = `cases (100k)`)) +
  geom_bar(stat = "identity", alpha = .85) +
  facet_wrap(~series, ncol = 2, scales = "free") +
  theme_minimal() +
  labs(
    title = "Share of COVID-19 incidents: most affected (full sample)",
    subtitle = "",
    x = "", y = "",
    caption = paste0("Date recorded: ", date_today, ".")
  ) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  scale_fill_viridis_d() -> p1

ggplotly(p1, tooltip = c("x", "y", "cases"))
```

### Most Affected (last 5 days)

```{r most.affected.5days}
# Aggregate data last five days --------------
## Extract full sample top countries
df_fin %>%
  filter(date >= (ymd(date_today) - 5)) %>%
  group_by(country) %>%
  summarize_at(c("confirmed", "dead", "recovered", "remaining"), max) -> df_topX_past5

# Extract list of topX countries for each category for full sample
rbind(
  extract_topX(df_topX_past5, X = x, "confirmed") %>% select(country, series),
  extract_topX(df_topX_past5, X = x, "dead") %>% select(country, series),
  extract_topX(df_topX_past5, X = x, "recovered") %>% select(country, series),
  extract_topX(df_topX_past5 %>% filter(country != "US"), X = x, "remaining") %>% select(country, series)
) %>% as_tibble() -> topX_countries_past5

## Plot last-5-days sample top countries
rbind(
  extract_topX(df_topX_past5, X = x, "confirmed"),
  extract_topX(df_topX_past5, X = x, "dead"),
  extract_topX(df_topX_past5, X = x, "recovered"),
  extract_topX(df_topX_past5 %>% filter(country != "US"), X = x, "remaining")
) %>%
  ggplot(aes(x = country, y = `share (%)`, fill = country, cases = `cases (100k)`)) +
  geom_bar(stat = "identity", alpha = .85) +
  facet_wrap(~series, nrow = 2, scales = "free") +
  theme_minimal() +
  labs(
    title = "Share of COVID-19 incidents: most affected (last five days)",
    subtitle = "",
    x = "", y = "",
    caption = paste0("Date recorded: ", date_today, ".")
  ) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  scale_fill_viridis_d() -> p2

ggplotly(p2, tooltip = c("x", "y", "cases"))
```

### Country Table

```{r country.table}
# Country tables   --------------

## Custom table container
sketch <- htmltools::withTags(table(
  class = "display",
  thead(
    tr(
      th(rowspan = 2, "Country"),
      th(rowspan = 2, "Date"),
      th(colspan = 4, "Cases")
    ),
    tr(
      lapply(c("confirmed", "dead", "recovered", "remaining"), th)
    )
  )
))

## Table
df_fin %>%
  select(-c(pop, iso2c)) %>%
  filter(date >= (ymd(date_today) - 10)) %>%
  group_by(country) %>%
  arrange(country, desc(date)) %>%
  datatable(
    container = sketch,
    rownames = FALSE,
    caption = "Table 1: Infections over last five recorded days.",
    options = list(
      pageLength = 10
    )
  )
```   

Austria  
===================================== 

```{r download_AGES}
# Download AGES data  --------------
download.file("https://covid19-dashboard.ages.at/data/data.zip", file.path("./data/AGES/data.zip"))
unzip(list.files(path = "./data/AGES/", pattern = "*.zip", full.names = TRUE),
      exdir="./data/AGES")
```




Column {data-width=600, .tabset}
-------------------------------------

```{r transform.state}
# Load raw data into R   --------------
readr::read_delim("data/AGES/data/CovidFaelle_Timeline.csv", 
    ";", escape_double = FALSE, trim_ws = TRUE) %>%
  select(state = Bundesland, date = Time, pop = AnzEinwohner, 
         cases = AnzahlFaelle, casesCum = AnzahlFaelleSum, cases7 = AnzahlFaelle7Tage,
         incidence = SiebenTageInzidenzFaelle,
         dead = AnzahlTotTaeglich, deadCum = AnzahlTotSum,
         recovered = AnzahlGeheiltTaeglich , recoveredCum = AnzahlGeheiltSum) %>%
  mutate(date = as.Date(substr(date, 1, 10), format = '%d.%m.%Y')) %>%
  select(state, date, pop, cases = casesCum, dead = deadCum, recovered = recoveredCum) %>%
  mutate(remaining = cases - dead - recovered) %>%
  filter(state != 'Österreich') -> df_states
```


### Instances by state (per thousand inhabitants)


```{r instances.state}
## Colour pattern
cp_custom <- c('#21908CFF', '#FDE725FF', '#440154FF')

df_states %>%
  mutate_at(.vars = c('cases', 'dead', 'recovered', 'remaining'), ~ round(./pop *1000, 2)) %>%
  pivot_longer(cols = 'dead':'remaining') %>%
  filter(value >0) -> df_plot

df_plot$name <- factor(df_plot$name, levels = c('recovered', 'dead', 'remaining'))

  ggplot(df_plot,
         aes(x = date, y = value, col = name)) +
  geom_line()  +
  facet_wrap(~state, ncol = 3)+
   theme_minimal() +
  labs(
    title = "COVID-19 incidents by state",
    subtitle = "",
    x = "", y = "",
    caption = paste0("Date recorded: ", date_today, ".")
  ) +
  theme(
    legend.position = "none"
  )   +
  scale_color_manual(values = cp_custom)  +
  scale_x_date(date_labels = "%e-%m") -> p3
  
  ggplotly(p3, tooltip = c('name','col', "y")) %>%   
     layout(hovermode = "x")
```






Column {data-width=400, .tabset}
-------------------------------------

```{r transform_district}
# Load raw data into R   --------------
readr::read_delim("data/AGES/data/CovidFaelle_Timeline_GKZ.csv", 
    ";", escape_double = FALSE, trim_ws = TRUE) %>%
  select(district = Bezirk, date = Time, pop = AnzEinwohner, 
         cases = AnzahlFaelle, casesCum = AnzahlFaelleSum, cases7 = AnzahlFaelle7Tage,
         incidence = SiebenTageInzidenzFaelle,
         dead = AnzahlTotTaeglich, deadCum = AnzahlTotSum,
         recovered = AnzahlGeheiltTaeglich , recoveredCum = AnzahlGeheiltSum) %>%
  mutate(date = as.Date(substr(date, 1, 10), format = '%d.%m.%Y')) %>%
  select(district, date, pop, cases = casesCum, dead = deadCum, recovered = recoveredCum) %>%
  mutate(remaining = cases - dead - recovered) -> df_district
```


### District Table

```{r district.table}
# District table   --------------

## Custom table container
sketch <- htmltools::withTags(table(
  class = "display",
  thead(
    tr(
      th(rowspan = 2, "District"),
      th(rowspan = 2, "Date"),
      th(colspan = 4, "Cases")
    ),
    tr(
      lapply(c("confirmed", "dead", "recovered", "remaining"), th)
    )
  )
))

## Table
df_district %>%
  select(-pop) %>%
  filter(date >= (ymd(date_today) - 10)) %>%
  group_by(district) %>%
  arrange(district, desc(date)) %>%
  datatable(
    container = sketch,
    rownames = FALSE,
    caption = "Table 1: Infections over last five recorded days.",
    options = list(
      pageLength = 10
    )
  )
```   



About {data-icon="fa-info-circle"}
=====================================

**COVID-19 Dashboard** 

[![View on Github](github.svg)](https://github.com/amannj/COVID-19)

- Kindly report any issues you may encounter to the [designated GitHub page](https://github.com/amannj/COVID-19/issues). Thank you.
- Created by [(c) 2021 by Juergen Amann](https://amannj.github.io).

--------------

**Data:**

- Data for cross-country analysis taken from the [COVID-19 (2019-nCoV) Data Repository by Johns Hopkins CSSE](https://github.com/CSSEGISandData/COVID-19). 
- Austrian data taken from [BMSGPK](https://www.sozialministerium.at/Informationen-zum-Coronavirus/Neuartiges-Coronavirus-(2019-nCov).html), [Gesundheitsministerium.at](https://info.gesundheitsministerium.at/) and
[derStandard](https://www.derstandard.at/story/2000115810293/aktuelle-zahlen-zum-coronavirus?ref=cta_red_s1). Population by district from [Statistik Austria](https://www.statistik.at/web_de/klassifikationen/regionale_gliederungen/politische_bezirke/index.html).
- Population data extracted from the [World Development Indicators (WDI) data base](https://databank.worldbank.org/reports.aspx?source=World-Development-Indicators) using package [WDI](https://github.com/vincentarelbundock/WDI).

**Shapefile Austria:**

- Shapefile for Austria taken from [EEA](https://www.eea.europa.eu/data-and-maps/data/eea-reference-grids-2/gis-files/austria-shapefile).
- Shapefile was modified using [QGIS](https://www.qgis.org/en/site/) to incorporate the most recent district re-classification in which district 'Wien Umgebung' was resolved; [click here for more information](https://en.wikipedia.org/wiki/Wien-Umgebung_District).

**Favicon:**

- Favicon taken from https://icons8.com/.

**Acknowledgements:** 

- Inspiration: Yanchang Zhao, COVID-19 Data Analysis with Tidyverse and Ggplot2 – Worldwide. RDataMining.com, 2020. URL: [http://www.rdatamining.com/docs/Coronavirus-data-analysis-world.pdf](http://www.rdatamining.com/docs/Coronavirus-data-analysis-world.pdf).

- Data preparation using the [Tidyverse](https://www.tidyverse.org/). Visualisation with [flexdashboard](https://rmarkdown.rstudio.com/flexdashboard/index.html).
